Bayesian statistics for repeated measures

### 2020-10-19 Manuscripts/Manuscript_AM_v1.Rmd
Generated first pdf!

---
title: "Bayesian statistics for repeated measures"
author: "Ariel Mundo, John Tipton, Timothy Muldoon"
bibliography: refs.bib
output:
  pdf_document: default
  html_document: default
  css: style.css
  word_document:
    reference_docx: docx_template.docx
mainfont: Times New Roman
link-citations: yes
subtitle: Their application and use in biomedical research
biblio-style: apalike
---

@@

Biomedical researchers typically employ a $frequentist$ approach to analyze longitudinal data. Such type of analysis is a hypothesis test using the $analysis of variance over repeated measures$ (repeated measures ANOVA or rm-ANOVA). Although the usual approach, rm-ANOVA is subject to multiple conditions to be valid, some of which are not easily verifiable. Moreover, this approach restricts the inferences it can extract from a longitudinal study, particularly when the data does not follow a linear trend (Figure 1. with line and "wiggly plot"), due to the inherent nature of the model and its limitations .Examples of such type of data are found in studies of tumor response to radio/chemotherapy in multiple models and clinical settings [[@vishwanath2009]]. Therefore, this study has three goals: a) Present in an amenable and practical manner the requisities  of a $frequentist$ approach over longitudinal data and how these limit the analysis b) present a different area of statistical analysis for longitudinal biomedical data $Bayesian Statistics$, which does not use *p-values*, and is less restrictive thereby allowing  a statical analysis that is based on the data itself. c)Iimplement b) over a set of simulated data that matches previously reported trends in longitudinal biomedical studies. With an emphasis on reproducibility by providing the code and dataset used, this will provide biomedical researchers a clear view of the advantages of Bayesian statistics for the analysis of longitudinal data.

>>>>

---
title: "Bayesian statistics for repeated measures"
author: "Ariel Mundo, John Tipton, Timothy Muldoon"
bibliography: refs.bib
output:
  pdf_document: default
  word_document:
    reference_docx: docx_template.docx
  css: style.css
  html_document: default
mainfont: Times New Roman
link-citations: yes
subtitle: Their application and use in biomedical research
biblio-style: apalike
---

@@

Biomedical researchers typically employ a $frequentist$ approach to analyze longitudinal data. Such type of analysis is a hypothesis test using the _analysis of variance over repeated measures_ (repeated measures ANOVA or rm-ANOVA). Although the usual approach, rm-ANOVA is subject to multiple conditions to be valid, some of which are not easily verifiable. Moreover, this approach restricts the inferences it can extract from a longitudinal study, particularly when the data does not follow a linear trend (Figure 1. with line and "wiggly plot"), due to the inherent nature of the model and its limitations .Examples of such type of data are found in studies of tumor response to radio/chemotherapy in multiple models and clinical settings [[@vishwanath2009]]. Therefore, this study has three goals: a) Present in an amenable and practical manner the requisities  of a $frequentist$ approach over longitudinal data and how these limit the analysis b) present a different area of statistical analysis for longitudinal biomedical data $Bayesian Statistics$, which does not use *p-values*, and is less restrictive thereby allowing  a statical analysis that is based on the data itself. c)Iimplement b) over a set of simulated data that matches previously reported trends in longitudinal biomedical studies. With an emphasis on reproducibility by providing the code and dataset used, this will provide biomedical researchers a clear view of the advantages of Bayesian statistics for the analysis of longitudinal data.
